### 2023-12-22 manuscript_PPC_AW(NHB).qmd

---
title: A Descriptive Analysis of How Psychology Journals Handle Post-Publication
  Critiques
format:
  pdf: default
  journal: sn-nature-human-behaviour
  keep-tex: true
author:
  - name: Annie Whamond
    affiliations:
      - name: University of Melbounre
        id: 1
        department: Melbourne School of Psychological Sciences
        address: 161 Barry St
        city: Carlton
        state: VIC
        postal-code: 3010
    email: whamonda@unimelb.edu.au
    corresponding: true
  - name: Simine Vazire
    affiliations:
      - ref: 1
  - name: Beth Clarke
    affiliations:
      - ref: 1
  - name: Nicholas Moodie
    affiliations:
      - ref: 1
  - name: Sarah Schiavone
    affiliations:
      - ref: 1
  - name: Robert T. Thibault
    affiliations:
      - name: Stanford University
        id: 2
        department: Meta Research Innovation Centre
        address: 450 Jane Stanford Way
        city: Stanford
        state: CA
        postal-code: 94305
  - name: Tom E. Hardwicke
    affiliations:
      - ref: 1
abstract: >
  Fostering critical discussion is essential to upholding the ethos of
  science as an unbiased, collaborative, and self-correcting endeavor. Journals
  can support this by offering options for post-publication critique, such as
  commentaries or letters, allowing peers to voice substantiated concerns of
  empirical articles. We provide a descriptive, cross-sectional analysis of
  post-publication critique policies and prevalence across both prominent
  randomly-selected and psychology journals. We found that, overall, the
  majority of psychology journals do not offer explicit options for
  post-publication critique. When they did, length and time-to-submit limits
  were often not stated. Critique prevalence was rare with only a single event
  identified. Given the widespread recognition of the replication crisis in
  psychology, allowing avenues for criticisms to be voiced may help increase the
  credibility within the field. To support transparency and self-correction,
  psychology journals should offer clearly stated post-publication critique
  options and encourage their use.
keywords:
  - metascience
  - post-publication cretique
  - psychology
  - credibility
bibliography: bibliography.bib

---
null

# Introduction {#sec-intro}

The Introduction section, of referenced text @greenwade93 expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.

Springer Nature does not impose a strict layout as standard however authors are advised to check the individual requirements for the journal they are planning to submit to as there may be journal-level preferences. When preparing your text please also be aware that some stylistic choices are not supported in full text XML (publication version), including coloured font. These will not be replicated in the typeset article if it is accepted.

# Using this template {#sec-template}

This Quarto template is unofficial and built out of Springer Nature's template. Some examples of commonly used commands and features are listed below, to help you get started.

As seen below, you can mix markdown and Latex with each other, though it's likely best to mostly use markdown.

## Cross Referencing

To reference a figure with example label "plot", use `@fig-plot`. Analogously, to reference a table with example label "data", use `@tbl-data`. To reference a section, such as the Introduction (@sec-intro), use `@sec-intro`.

For complete information on cross referencing with Quarto, see <https://quarto.org/docs/authoring/cross-references.html>.

## Citations

Quarto formats citations and references automatically using the bibliography records in your .bib file. For a citation in parentheses use [@greenwade93] and for a text citation: @greenwade93. Multiple citations can be given as [@greenwade93; @knuth1984texbook]. Drop the author like so [-@greenwade93].

# Tables and Figures

To include figures, you can use Quarto syntax.

![An example figure (an empty plot)](fig.eps){#fig-plot}

For both figures and tables, you can use LaTeX syntax if you need additional customizability.

For example, to use footnotes within a table, you may want to use LaTeX.

```{=tex}
\begin{table}[h]
\caption{LaTeX caption text}\label{tbl-data}%
\begin{tabular}{@{}llll@{}}
\toprule
Column 1 & Column 2  & Column 3 & Column 4\\
\midrule
row 1    & data 1   & data 2  & data 3  \\
row 2    & data 4   & data 5\footnotemark[1]  & data 6  \\
row 3    & data 7   & data 8  & data 9\footnotemark[2]  \\
\botrule
\end{tabular}
\footnotetext{Source: This is an example of table footnote. This is an example of table footnote.}
\footnotetext[1]{Example for a first table footnote. This is an example of table footnote.}
\footnotetext[2]{Example for a second table footnote. This is an example of table footnote.}
\end{table}
```
# References

<!-- the pattern below controls the placement of the references -->

::: {#refs}
:::