### 2024-01-05 manuscript/manuscript_NHB.Rmd
manuscript draft
updating changes

---
title: "A Descriptive Analysis of How Psychology Journals Handle Post-Publication
  Critiques"
author: Whamond, A., Vazire, S., Clarke, B., Moodie, N., Schiavone, S., Thibault,
  R. T., & Hardwicke, T. E.
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  word_document: default
abstract: "Fostering critical discussion is essential to upholding the ethos of science
  as an unbiased, collaborative, and self-correcting endeavor. Journals can support
  this by offering options for post-publication critique, such as commentaries or
  letters, allowing peers to voice substantiated concerns of empirical articles. We
  provide a descriptive, cross-sectional analysis of post-publication critique policies
  and prevalence across both prominent randomly-selected and psychology journals.
  We found that 38% of prominent and 23% of randomly-selected psychology journals
  offered explicit options for post-publication critique. Length and time-to-submit
  limits were often not stated. Critique prevalence was rare with only a single event
  identified among two samples of 101 articles. Given the widespread recognition of
  the replication crisis in psychology, allowing avenues for criticisms to be voiced
  may help increase the credibility within the field. To support transparency and
  self-correction, psychology journals should offer clearly stated post-publication
  critique options and encourage their use.\n"
keywords: "post-publication critique, peer review, psychology, meta-research, journal
  policy, self-correction, scientific criticism"

bibliography: references.bib
link-citations: true
csl: nature.csl

floatsintext: false
figurelist: false
tablelist: false
footnotelist: false
linenumbers: true
mask: true
draft: true
always_allow_html: true
urlcolor: blue
---

```{r setup, include = FALSE}

library(papaja) # article template
library(tidyverse)
library(ggplot2) # figure creation
library(ggrepel) # for ggplot text labels
library(patchwork) # for multipanel plot layouts
library(RColorBrewer) # data visualisation
library(dplyr)
library(kableExtra) # for tables
library(knitr) # for literate programming
library(janitor) # for data munging
library(tidylog) # for inline code feedback
library(here) # for finding files
library(binom) # for calculating CIs
library(citation) # for organizing citations
library(tinytex)
```

```{r echo=FALSE, message=FALSE}
# Load processed data files
#'ProjCode', 'GitClone', 'PROJ_2023_psych-journal-critique',
d_articles <- read_csv(here('data', 'processed', 'd_articles.csv'), show_col_types = F) # article data 
d_journals <- read_csv(here('data', 'processed', 'd_journals.csv'), show_col_types = F) # journal data

```

```{r echo=FALSE, message=FALSE}
# Create separate tibbles for Prominent and Randomly-selected journal samples

dj_prominent <- d_journals %>%
  filter(sample_id != "random") %>% # remove entries from the random sample
  filter(is_empirical == TRUE) # remove ineligible (non-empirical) journals

dj_random <- d_journals %>%
  filter(sample_id != "prominent") %>% # remove entries from the prominent sample
  filter(is_empirical == TRUE) # remove ineligible (non-empirical) journals
```

# Main
*Add 'Introduction' section*

```{r echo = FALSE, message = FALSE}
# Create Table for describing the different common PPC types and restrictions (based on Hardwicke et al., 2022 findings)
Type <- c("Letters", "Commentaries", "Web Comments")
MedianLength <- c("500", "1200", "600")
MedianTime <- c("8", "26", "NA")
MedianRefs <- c("6", "15", "8")
PeerReview <- c("20%", "55%", "0%")
DOIs <- c("Yes", "Yes", "No")
Utility <- c("Concise criticisms submitted shortly after target article published", "More in-depth critical reviews", "Flexible length with almost immediate visibility to readership")

PPCtypes <- data.frame(Type, MedianLength, MedianTime, MedianRefs, PeerReview, DOIs, Utility)
```

``` {r echo = FALSE, results = 'asis'}
PPCtypes %>%
  kbl(align = c(rep("l", 1), rep("c", 5), rep("l", 1)), caption = "Common post-publication critique types including median restrictions, peer review policies, and descriptions",
      col.names = c("Type", "Length", "Time-to-submit", "References", "Peer-review", "DOI", "Description")) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, align = "c")

```

# Results

## Part 1: Post-Publication Critique Policies

For descriptive statistics on journal subdisciplines, post-publication critique acceptance, and COPE membership for both prominent and randomly-selected samples, see Supplementary Information ##. A full list of the journals included in this study, along with their post-publication critique options and restrictions, is provided in Supplementary Information ##.

## Prominent Psychology Journals

```{r prominent journal-characteristics, echo=FALSE, message=FALSE}

# Ascertain prominent journal JIF characteristics

promJournalCharacteristics <- dj_prominent %>% 
  filter(ppc_type == "A") %>%       # Keep only single entry for each journal
  summarise(promMedian_jif = round(median(jif),2), 
            promMin_jif = round(min(jif),2),
            promMax_jif = round(max(jif),2),
            promIQR_jif = round(IQR(jif),2)
  )

```

In our sample of 100 prominent journals, the median 2021 Journal Impact Factor was `r promJournalCharacteristics %>% pull(promMedian_jif)` (IQR = `r promJournalCharacteristics %>% pull(promIQR_jif)`, range = `r promJournalCharacteristics %>% pull(promMin_jif)` - `r promJournalCharacteristics %>% pull(promMax_jif)`). As this represents the entire population of interest for this section of the study, confidence intervals (CIs) are not included for this section of the study. Figure *XX* shows the observed journal policies and, of the explicitly offered critique options identified, how restrictions on length and time-to-submit were stated.

### *Post-Publication Critique Policy and Restriction Statements in Prominent Psychology Journals*

```{r, echo = FALSE, message=FALSE}
# Filter and transform restrictions for ppc_length
prominent_length_data <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  filter(has_ppc != "ARCHIVE") %>% # removes 'Archive Only' PPC types
  mutate(ppc_length = case_when(
    ppc_length %in% c(1000, 1100, 1200, 1500, 2000, 2500, 2650, 3000, 3500, 400, 500, 5000, 750, 800, 8000) ~ 'Quantitative',
    ppc_length == "NOT STATED" ~ "Not Stated",
    TRUE ~ 'Qualitative' # Summarises all length limits as 'Quantitative', 'Qualitative', or 'Not Stated'
  )) %>% 
  group_by(ppc_length) %>%
  summarise(count = n())

# Filter and transform restrictions for ppc_time
prominent_time_data <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  filter(has_ppc != "ARCHIVE") %>% # removes 'Archive Only' PPC types
  mutate(ppc_time = case_when(
    ppc_time %in% c(13.05, 26.1, 39.15, 4, 52) ~ 'Quantitative',
    ppc_time == "NOT STATED" ~ "Not Stated",
    TRUE ~ 'Qualitative' # Summarises all time limits as 'Quantitative', 'Qualitative', or 'Not Stated'
  )) %>%
  group_by(ppc_time) %>%
  summarise(count = n())

# Combine the data and unite the columns
combined_data <- bind_rows(
  prominent_length_data %>% mutate(Code = "Length"),
  prominent_time_data %>% mutate(Code = "Time-to-submit")
) %>%
unite(limits, ppc_length, ppc_time) %>%
  mutate(limits = case_when(limits == "Not Stated_NA" ~ "Not Stated",
                   limits == "Qualitative_NA" ~ "Qualitative",
                   limits == "Quantitative_NA" ~ "Quantitative",
                   limits == "NA_Not Stated" ~ "Not Stated",
                   limits == "NA_Qualitative" ~ "Qualitative",
                   limits == "NA_Quantitative" ~ "Quantitative",
                   TRUE ~ as.character(limits)
                   ))

# Create the first graph
graph1 <- ggplot(combined_data, aes(x = Code, y = count, fill = limits)) +
  geom_bar(stat = "identity", 
           position = "stack",
           colour = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Blues") +
  labs(x = "Restriction Type", y = "Count", fill = "Restrictions")

# Create the second graph
graph2 <- dj_prominent %>%
  filter(ppc_type == "A") %>% # Keeps only first entry for each journal
  filter(!is.na(has_ppc)) %>% # Remove journal rows with no (NA) PPC options 
  mutate(has_ppc = case_when(
    has_ppc == 'ARCHIVE' ~ 'Yes Implicit',
    has_ppc == 'NO EXPLICIT' ~ 'No Explicit',
    has_ppc == 'NO IMPLICIT' ~ 'No Implicit',
    has_ppc == 'YES' ~ 'Yes Explicit',
    TRUE ~ has_ppc
  )) %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = "", y = count, fill = has_ppc)) +
  geom_bar(stat = "identity",
           colour = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Oranges") +
  labs(x = "Journal Policy", y = "Count", fill = "Accepts PPC")

# Arrange both graphs in the same figure
combined_plots <- graph2 + graph1
combined_plots

```

*Note.* PPC = post-publication critique. Orange bar (left) indicates policy statements for each journal included in sample (n = 100). Blue bars (right) indicate how length and time-to-submit restrictions were stated in cases where journals had explicit post-publication critique policies in place (n = 41). Implicit acceptance denoted by archival examples of critiques despite absence of stated policy.

```{r echo=FALSE, message=FALSE}
# Length restriction statement types of explicit PPC types (prominent sample)

# Calculate the count and percentage for "Quantitative"
prom_quantitative_length <- prominent_length_data %>%
 filter(ppc_length == "Quantitative") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_length_data$count) * 100)

# Calculate the count and percentage for "Qualitative"
prom_qualitative_length <- prominent_length_data %>%
 filter(ppc_length == "Qualitative") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_length_data$count) * 100)

# Calculate the count and percentage for "Not Stated"
prom_notStated_length <- prominent_length_data %>%
 filter(ppc_length == "Not Stated") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_length_data$count) * 100)

# Range
prom_length_range <- dj_prominent %>%
  filter(!is.na(ppc_length)) %>%
  filter(ppc_length != "Qualitative") %>%
  filter(ppc_length != "NOT STATED") %>%
  mutate(ppc_length = as.numeric(ppc_length)) %>%
  summarise(prom_median_length = median(ppc_length),
            prom_min_length = min(ppc_length),
            prom_max_length = max(ppc_length)
            )

# Time-to-submit restriction statement types of explicit PPC types (prominent sample)

# Calculate the count and percentage for "Quantitative"
prom_quantitative_time <- prominent_time_data %>%
 filter(ppc_time == "Quantitative") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_time_data$count) * 100)

# Calculate the count and percentage for "Qualitative"
prom_qualitative_time <- prominent_time_data %>%
 filter(ppc_time == "Qualitative") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_time_data$count) * 100)

# Calculate the count and percentage for "Not Stated"
prom_notStated_time <- prominent_time_data %>%
 filter(ppc_time == "Not Stated") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_time_data$count) * 100)

# Range
prom_time_range <- dj_prominent %>%
  filter(!is.na(ppc_time)) %>%
  filter(ppc_time != "Qualitative") %>%
  filter(ppc_time != "NOT STATED") %>%
  mutate(ppc_time = as.numeric(ppc_time)) %>%
  summarise(prom_median_time = median(ppc_time),
            prom_min_time = min(ppc_time),
            prom_max_time = max(ppc_time)
            )

# Reference restriction statement types of explicit PPC types (prominent sample)
prominent_ref_data <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>% # keep only entries that identify PPC option
  filter(has_ppc != "ARCHIVE") %>% # remove implicit yes examples
  mutate(ppc_ref = case_when(
    ppc_ref %in% c(10, 15, 20, 5) ~ 'Quantitative',
    ppc_ref == "NOT STATED" ~ "Not Stated",
    TRUE ~ 'Qualitative' # Summarises all time limits as 'Quantitative', 'Qualitative', or 'Not Stated'
  )) %>%
  group_by(ppc_ref) %>%
  summarise(count = n()) %>%
  ungroup()

# Calculate the count and percentage for "Quantitative"
prom_quantitative_ref <- prominent_ref_data %>%
 filter(ppc_ref == "Quantitative") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_ref_data$count) * 100)

# Calculate the count and percentage for "Not Stated"
prom_notStated_ref <- prominent_ref_data %>%
 filter(ppc_ref == "Not Stated") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_ref_data$count) * 100)

# Range
prom_ref_range <- dj_prominent %>%
  filter(!is.na(ppc_ref)) %>%
  filter(ppc_ref != "Qualitative") %>%
  filter(ppc_ref != "NOT STATED") %>%
  mutate(ppc_ref = as.numeric(ppc_ref)) %>%
  summarise(prom_median_ref = median(ppc_ref),
            prom_min_ref = min(ppc_ref),
            prom_max_ref = max(ppc_ref)
            )

# Peer-review statements of explicit PPC types (prominent sample)
prominent_review_data <- dj_prominent %>%
  filter(!is.na(ppc_review)) %>% # keep only entries that identify PPC option
  filter(has_ppc != "ARCHIVE") %>% # remove implicit yes examples
  mutate(ppc_review = case_when(
    ppc_review == 'YES' ~ 'Yes',
    ppc_review == 'NO' ~ 'No',
    ppc_review == "NOT STATED" ~ "Not Stated")
  ) %>%
  group_by(ppc_review) %>%
  summarise(count = n()) %>%
  ungroup()

# Calculate the count and percentage for "Yes"
prom_yes_review <- prominent_review_data %>%
 filter(ppc_review == "Yes") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_review_data$count) * 100)

# Calculate the count and percentage for "No"
prom_no_review <- prominent_review_data %>%
 filter(ppc_review == "No") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_review_data$count) * 100)

# Calculate the count and percentage for "Not Stated"
prom_notStated_review <- prominent_review_data %>%
 filter(ppc_review == "Not Stated") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(prominent_review_data$count) * 100)

# Calculate the count and percentage for 2 types
prominent2types <- dj_prominent %>%
  filter(ppc_type == "B") %>%
  filter(!is.na(ppc_name)) %>%
  count()

```

As shown in Figure *XX*, post-publication critique policies were explicitly stated by `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "YES") / n() * 100, "%"))` of journals, and we inferred that they were implicitly accepted (as denoted by archival examples) in `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "ARCHIVE") / n() * 100, "%"))` of journals. `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "NO IMPLICIT") / n() * 100, "%"))` had no advertised post-publication critique policy, and `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "NO EXPLICIT") / n() * 100, "%"))` of journals explicitly stated that they did not accept critiques (*Clinical Psychology: Science and Practice and Psychological Bulletin*). `r paste0(prominent2types, "%")` of journals accepted two types (commentaries and letters); *International Journal of Mental Health and Addiction*, *Journal of Environmental Psychology*, and *Psychological Science*.

Among the 41 explicitly stated post-publication critique options, restrictions on length were quantitatively stated (e.g., '1,000 words') in `r paste(prom_quantitative_length$count, "(", round(prom_quantitative_length$percentage), "%)", sep = "")` cases, qualitatively stated (e.g. 'brief') in  `r paste(prom_qualitative_length$count, "(", round(prom_qualitative_length$percentage), "%)", sep = "")` cases, and not stated in `r paste(prom_notStated_length$count, "(", round(prom_notStated_length$percentage), "%)", sep = "")` cases. Time-to-submit was quantitatively stated (e.g., '4 weeks') in `r paste(prom_quantitative_time$count, "(", round(prom_quantitative_time$percentage), "%)", sep = "")` cases, qualitatively stated (e.g., 'timely') in `r paste(prom_qualitative_time$count, "(", round(prom_qualitative_time$percentage), "%)", sep = "")` cases, and not stated in `r paste(prom_notStated_time$count, "(", round(prom_notStated_time$percentage), "%)", sep = "")` cases. Number of references allowed was quantitatively stated (e.g., 'max 20') in `r paste(prom_quantitative_ref$count, "(", round(prom_quantitative_ref$percentage), "%)", sep = "")` cases and not stated in `r paste(prom_notStated_ref$count, "(", round(prom_notStated_ref$percentage), "%)", sep = "")`. Whether critiques were sent for external peer-review was only stated in `r paste((prom_yes_review$count + prom_no_review$count), "(", round(prom_yes_review$percentage + prom_no_review$percentage), "%)", sep = "")` cases, in `r paste(prom_yes_review$count, "(", round(prom_yes_review$percentage), "%)", sep = "")` cases this was in the affirmative.

When stated quantitatively, length limits ranged from `r prom_length_range %>% pull(prom_min_length)` - `r prom_length_range %>% pull(prom_max_length)` words, time-to-submit limits ranged from `r prom_time_range %>% pull(prom_min_time)` - `r prom_time_range %>% pull(prom_max_time)` weeks, reference limits ranged between `r prom_ref_range %>% pull(prom_min_ref)` -`r prom_ref_range %>% pull(prom_max_ref)`. Table *XX* shows descriptive statistics for stated quantitative restrictions, broken down by critique type.

### Table *XX*

As expected, results in Table 1 show that commentaries were generally less restrictive on length with the median word limit being 2.5 times that of letters. The median number of references allowed was 1.5 times greater in commentaries compared to letters. Commentaries were also 3.3 times more likely to be sent for peer-review. Interestingly, time-to-submit was less restrictive for letters than commentaries, however, this was only reported in two and four cases (respectively) so limited inferences can be made from this.

## Randomly-selected Psychology Journals
```{r randomly-selected journal-characteristics, echo=FALSE, message=FALSE}

# Ascertain randonly-selected journal JIF characteristics

randJournalCharacteristics <- dj_random %>% 
  filter(ppc_type == "A") %>%
  filter(!is.na(jif)) %>%
  summarise(randMedian_jif = round(median(jif),2),
            randMin_jif = round(min(jif),2),
            randMax_jif = round(max(jif),2),
            randIQR_jif = round(IQR(jif),2)
  )

```
Among our sample of 100 randomly-selected psychology journals, Journal Impact Factor data was unavailable for 25 journals. Median 2021 Journal Impact Factor was `r randJournalCharacteristics %>% pull(randMedian_jif)` (IQR = `r randJournalCharacteristics %>% pull(randIQR_jif)`, range = `r randJournalCharacteristics %>% pull(randMin_jif)` - `r randJournalCharacteristics %>% pull(randMax_jif)`). Figure *XX* shows the observed journal policies and, of the explicitly offered critique options identified, how restrictions on length and time-to-submit were stated.

### Figure *XX*: *Post-Publication Critique Policy and Restriction Statements in Randomly-selected Psychology Journals*
```{r echo=FALSE, message=FALSE}
# Create and combine restriction statement graphs with PPC acceptance (random sample)

# Filter and process data for ppc_length
random_length_data <- d_journals %>%
  filter(sample_id %in% c("random", "both")) %>%
  filter(!is.na(ppc_name)) %>%
  filter(has_ppc != "ARCHIVE") %>%
  mutate(ppc_length = case_when(
    ppc_length %in% c(1000, 1100, 1200, 1500, 2000, 2500, 2650, 3000, 3500, 400, 500, 5000, 750, 800, 8000) ~ 'Quantitative',
    ppc_length == "NOT STATED" ~ "Not Stated",
    TRUE ~ 'Qualitative'
  )) %>%
  group_by(ppc_length) %>%
  summarise(count = n())

# Filter and process data for ppc_time
random_time_data <- d_journals %>%
  filter(sample_id %in% c("random", "both")) %>%
  filter(!is.na(ppc_name)) %>%
  filter(has_ppc != "ARCHIVE") %>%
  mutate(ppc_time = case_when(
    ppc_time %in% c(13.05, 26.1, 39.15, 4, 52) ~ 'Quantitative',
    ppc_time == "NOT STATED" ~ "Not Stated",
    TRUE ~ 'Qualitative'
  )) %>%
  group_by(ppc_time) %>%
  summarise(count = n())

# Combine the data and unite the columns
combined_data_random <- bind_rows(
  random_length_data %>% mutate(Code = "Length"),
  random_time_data %>% mutate(Code = "Time-to-submit")
) %>%
unite(limits, ppc_length, ppc_time) %>%
  mutate(limits = case_when(limits == "Not Stated_NA" ~ "Not Stated",
                   limits == "Qualitative_NA" ~ "Qualitative",
                   limits == "Quantitative_NA" ~ "Quantitative",
                   limits == "NA_Not Stated" ~ "Not Stated",
                   limits == "NA_Qualitative" ~ "Qualitative",
                   limits == "NA_Quantitative" ~ "Quantitative",
                   TRUE ~ as.character(limits)
                   ))

# Create the first graph
graph4 <- ggplot(combined_data_random, aes(x = Code, y = count, fill = limits)) +
  geom_bar(stat = "identity", 
           position = "stack",
           colour = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Blues") +
  labs(x = "Restriction Type", y = "Count", fill = "Restrictions")

# Create the second graph
graph3 <- d_journals %>%
  filter(sample_id == "random" | sample_id == "both") %>%
  filter(ppc_type == "A") %>%
  filter(!is.na(has_ppc)) %>%
  mutate(has_ppc = case_when(
    has_ppc == 'ARCHIVE' ~ 'Yes Implicit',
    has_ppc == 'NO EXPLICIT' ~ 'No Explicit',
    has_ppc == 'NO IMPLICIT' ~ 'No Implicit',
    has_ppc == 'YES' ~ 'Yes Explicit',
    TRUE ~ has_ppc
  )) %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = "", y = count, fill = has_ppc)) +
  geom_bar(stat = "identity",
           colour = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Oranges") +
  labs(x = "Journal Policy", y = "Count", fill = "Accepts PPC")

# Arrange both graphs in the same figure
combined_plots <- graph3 + graph4
combined_plots
```

``` {r echo = FALSE, message = FALSE}

# Calculate Wilson Confidence Intervals
sample_size <- 100
confidence_level <- 0.95

#CIs for 'Yes Explicit'
randomYes <- dj_random %>%
  filter(ppc_type == "A") %>%
  filter(has_ppc == "YES") %>%
  count()

ci_yesE <- binom.confint(randomYes$n, sample_size, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_yesE_LL <- round(ci_yesE$lower*100, 1)
ci_yesE_UL <- round(ci_yesE$upper*100, 1)

#CIs for 'Yes Implicit'
randomArchive <- dj_random %>%
  filter(ppc_type == "A") %>%
  filter(has_ppc == "ARCHIVE") %>%
  count()

ci_yesI <- binom.confint(randomArchive$n, sample_size, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_yesI_LL <- round(ci_yesI$lower*100, 1)
ci_yesI_UL <- round(ci_yesI$upper*100, 1)

#CIs for 'No Implicit'
randomNoImp <- dj_random %>%
  filter(ppc_type == "A") %>%
  filter(has_ppc == "NO IMPLICIT") %>%
  count()

ci_noI <- binom.confint(randomNoImp$n, sample_size, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_noI_LL <- round(ci_noI$lower*100, 1)
ci_noI_UL <- round(ci_noI$upper*100, 1)

#CIs for 'No Explicit'
randomNoExp <- dj_random %>%
  filter(ppc_type == "A") %>%
  filter(has_ppc == "NO EXPLICIT") %>%
  count()

ci_noE <- binom.confint(randomNoExp$n, sample_size, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_noE_LL <- round(ci_noE$lower*100, 1)
ci_noE_UL <- round(ci_noE$upper*100, 1)

# CIs for 2 types
random2types <- dj_random %>%
  filter(ppc_type == "B") %>%
  filter(!is.na(ppc_name)) %>%
  count()

ci_2typ <- binom.confint(random2types$n, sample_size, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_2typ_LL <- round(ci_2typ$lower*100, 1)
ci_2typ_UL <- round(ci_2typ$upper*100, 1)
```

*Note.* PPC = post-publication critique. Orange bar (left) indicates policy statements for each journal included in sample (n = 100). Blue bars (right) indicate how length and time-to-submit restrictions were stated in cases where journals had explicit post-publication critique policies in place (n = 25).

```{r echo=FALSE, message=FALSE}
# Length restriction statement types of explicit PPC types (random sample)

# Calculate the count and percentage for "Quantitative"
rand_quantitative_length <- random_length_data %>%
 filter(ppc_length == "Quantitative") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_length_data$count) * 100)

# Calculate the count and percentage for "Qualitative"
rand_qualitative_length <- random_length_data %>%
 filter(ppc_length == "Qualitative") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_length_data$count) * 100)

# Calculate the count and percentage for "Not Stated"
rand_notStated_length <- random_length_data %>%
 filter(ppc_length == "Not Stated") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_length_data$count) * 100)

# Range
rand_length_range <- dj_random %>%
  filter(!is.na(ppc_length)) %>%
  filter(ppc_length != "Qualitative") %>%
  filter(ppc_length != "NOT STATED") %>%
  mutate(ppc_length = as.numeric(ppc_length)) %>%
  summarise(prom_median_length = median(ppc_length),
            rand_min_length = min(ppc_length),
            rand_max_length = max(ppc_length)
            )

# Time-to-submit restriction statement types of explicit PPC types (prominent sample)

# Calculate the count and percentage for "Quantitative"
rand_quantitative_time <- random_time_data %>%
 filter(ppc_time == "Quantitative") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_time_data$count) * 100)

# Calculate the count and percentage for "Qualitative"
rand_qualitative_time <- random_time_data %>%
 filter(ppc_time == "Qualitative") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_time_data$count) * 100)

# Calculate the count and percentage for "Not Stated"
rand_notStated_time <- random_time_data %>%
 filter(ppc_time == "Not Stated") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_time_data$count) * 100)

# Range
rand_time_range <- dj_random %>%
  filter(!is.na(ppc_time)) %>%
  filter(ppc_time != "Qualitative") %>%
  filter(ppc_time != "NOT STATED") %>%
  mutate(ppc_time = as.numeric(ppc_time)) %>%
  summarise(rand_median_time = median(ppc_time),
            rand_min_time = min(ppc_time),
            rand_max_time = max(ppc_time)
            )

# Reference restriction statement types of explicit PPC types (prominent sample)
random_ref_data <- dj_random %>%
  filter(!is.na(ppc_name)) %>% # keep only entries that identify PPC option
  filter(has_ppc != "ARCHIVE") %>% # remove implicit yes examples
  mutate(ppc_ref = case_when(
    ppc_ref %in% c(10, 15, 20, 5) ~ 'Quantitative',
    ppc_ref == "NOT STATED" ~ "Not Stated",
    TRUE ~ 'Qualitative' # Summarises all time limits as 'Quantitative', 'Qualitative', or 'Not Stated'
  )) %>%
  group_by(ppc_ref) %>%
  summarise(count = n()) %>%
  ungroup()

# Calculate the count and percentage for "Quantitative"
rand_quantitative_ref <- random_ref_data %>%
 filter(ppc_ref == "Quantitative") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_ref_data$count) * 100)

# Calculate the count and percentage for "Not Stated"
rand_notStated_ref <- random_ref_data %>%
 filter(ppc_ref == "Not Stated") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_ref_data$count) * 100)

# Range
rand_ref_range <- dj_random %>%
  filter(!is.na(ppc_ref)) %>%
  filter(ppc_ref != "Qualitative") %>%
  filter(ppc_ref != "NOT STATED") %>%
  mutate(ppc_ref = as.numeric(ppc_ref)) %>%
  summarise(rand_median_ref = median(ppc_ref),
            rand_min_ref = min(ppc_ref),
            rand_max_ref = max(ppc_ref)
            )

# Peer-review statements of explicit PPC types (prominent sample)
random_review_data <- dj_random %>%
  filter(!is.na(ppc_review)) %>% # keep only entries that identify PPC option
  filter(has_ppc != "ARCHIVE") %>% # remove implicit yes examples
  mutate(ppc_review = case_when(
    ppc_review == 'YES' ~ 'Yes',
    ppc_review == 'NO' ~ 'No',
    ppc_review == "NOT STATED" ~ "Not Stated"
  )) %>%
  group_by(ppc_review) %>%
  summarise(count = n()) %>%
  ungroup()

# Calculate the count and percentage for "Yes"
rand_yes_review <- random_review_data %>%
 filter(ppc_review == "Yes") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_review_data$count) * 100)

# Calculate the count and percentage for "No"
rand_no_review <- random_review_data %>%
 filter(ppc_review == "No") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_review_data$count) * 100)

# Calculate the count and percentage for "Not Stated"
rand_notStated_review <- random_review_data %>%
 filter(ppc_review == "Not Stated") %>%
 summarise(count = sum(count), percentage = sum(count) / sum(random_review_data$count) * 100)

```
As shown in Figure *XX*, post-publication critique policies were explicitly stated by `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "YES") / n() * 100, "%"))` of journals (`r paste0(round(confidence_level*100), "% CI [", ci_yesE_LL, "%, ", ci_yesE_UL, "%]")`), and we inferred that they were implicitly accepted (as denoted by archival examples) in `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "ARCHIVE") / n() * 100, "%"))` of journals (`r paste0(round(confidence_level*100), "% CI [", ci_yesI_LL, "%, ", ci_yesI_UL, "%]")`). `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "NO IMPLICIT") / n() * 100, "%"))` of journals (`r paste0(round(confidence_level*100), "% CI [", ci_noI_LL, "%, ", ci_noI_UL, "%]")`) had no advertised post-publication critique policy, and `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "NO EXPLICIT") / n() * 100, "%"))` of journals (`r paste0(round(confidence_level*100), "% CI [", ci_noE_LL, "%, ", ci_noE_UL, "%]")`) explicitly stated that they did not accept critiques (*Clinical Psychology: Science and Practice and Psychological Bulletin*, also included in sample for prominent journals). `r paste0(random2types, "%")` of journals accepted two types (`r paste0(round(confidence_level*100), "% CI [", ci_2typ_LL, "%, ", ci_2typ_UL, "%]")`); *Cortex* offered commentaries and 'verification reports' [@Chambers2020] and *Journal of Cognition and Culture* offered commentaries and letters.

Among the 25 cases of explicitly accepted post-publication critiques, restrictions on length were quantitatively stated in `r paste(rand_quantitative_length$count, "(", round(rand_quantitative_length$percentage), "%)", sep = "")` cases, and not stated in `r paste(rand_notStated_length$count, "(", round(rand_notStated_length$percentage), "%)", sep = "")` cases. Time-to-submit was quantitatively stated in `r paste(rand_quantitative_time$count, "(", round(rand_quantitative_time$percentage), "%)", sep = "")` case, and qualitatively stated in `r paste(rand_qualitative_time$count, "(", round(rand_qualitative_time$percentage), "%)", sep = "")`, and not stated in `r paste(rand_notStated_time$count, "(", round(rand_notStated_time$percentage), "%)", sep = "")` cases. Number of references allowed was stated quantitatively in `r paste(rand_quantitative_ref$count, "(", round(rand_quantitative_ref$percentage), "%)", sep = "")` cases, and not stated in `r paste(rand_notStated_ref$count, "(", round(rand_notStated_ref$percentage), "%)", sep = "")` cases. Whether critiques were peer-reviewed was stated in `r paste((rand_yes_review$count + rand_no_review$count), "(", round(rand_yes_review$percentage + rand_no_review$percentage), "%)", sep = "")` cases, in `r paste(rand_yes_review$count, "(", round(rand_yes_review$percentage), "%)", sep = "")` cases this was in the affirmative.

When stated quantitatively, length limits ranged from `r rand_length_range %>% pull(rand_min_length)` - `r rand_length_range %>% pull(rand_max_length)` words, time-to-submit limits were only stated by one journal (as 9 months), references ranged between `r rand_ref_range %>% pull(rand_min_ref)` - `r rand_ref_range %>% pull(rand_max_ref)`. Table *XX* shows descriptive statistics for these restrictions, broken down by critique type.

### Table *XX*

As expected, Table *XX* shows that commentaries were overall less restrictive on length with the median word limit being double that of letters. Commentaries were also three times more likely to state quantitative length restrictions than letters. Restrictions on time-to-submit, references, and whether critiques would be sent for peer-review were not stated for any policies on letters. Similarly, no details on restrictions or peer-review were stated for 'verification reports' (the one 'other' critique type included in this sample).

## Part 2: Post-Publication Critique Prevalence
```{r echo = FALSE, message=FALSE}
# Filter articles by sample and calculate CIs

# For articles randomly sampled from prominent psychology journals
da_prominent <- d_articles %>%
  filter(exclude == "RETAIN") %>% # remove excluded articles
  filter(sample_id == "prominent") # keep only articles from prominent journals

sample_size_a <- 101

# Linked PPCs
prominent_linked <- da_prominent %>% # counts number of linked PPC events
  filter(ppc_linked == "TRUE") %>%
  count()

ci_linked <- binom.confint(prominent_linked$n, sample_size_a, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_linked_LL <- round(ci_linked$lower*100, 1)
ci_linked_UL <- round(ci_linked$upper*100, 1)

# Instance PPCs
prominent_instance <- da_prominent %>% # counts number of linked PPC events
  filter(exclude_reason == "IS PPC") %>%
  count()

ci_instance <- binom.confint(prominent_instance$n, sample_size_a, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_instance_LL <- round(ci_instance$lower*100, 1)
ci_instance_UL <- round(ci_instance$upper*100, 1)

# For articles randomly sampled from all psychology journals
da_random <- d_articles %>%
  filter(exclude == "RETAIN") %>% # remove excluded articles
  filter(sample_id == "random") # keep only randomly-selected articles

# Linked PPCs
random_linked <- da_random %>%
  filter(ppc_linked == "TRUE") %>%
  count()

ci_linkedR <- binom.confint(random_linked$n, sample_size_a, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_linkedR_LL <- round(ci_linkedR$lower*100, 1)
ci_linkedR_UL <- round(ci_linkedR$upper*100, 1)

# Instance PPCs
random_instance <- da_random %>% # counts number of linked PPC events
  filter(exclude_reason == "IS PPC") %>%
  count()

ci_instanceR <- binom.confint(random_instance$n, sample_size_a, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_instanceR_LL <- round(ci_instanceR$lower*100, 1)
ci_instanceR_UL <- round(ci_instanceR$upper*100, 1)
```
**2a) Articles Published in Prominent Psychology Journals**

Prevalence was defined in two ways: linked critiques (links to critiques appearing on same webpage as target article) and instances of critiques (articles that are themselves critiques). Only one linked post-publication critique event was found within our sample of 101 articles published in prominent psychology journals, yielding a prevalence estimate of `r da_prominent %>% summarise(percentage = paste0(round(sum(as.integer(ppc_linked == "TRUE")) / n() * 100, 1), "%"))` (`r paste0(round(confidence_level*100), "% CI [", ci_linked_LL, "%, ", ci_linked_UL, "%]")`). No instances of post-publication critiques were found (0%, `r paste0(round(confidence_level*100), "% CI [", ci_instance_LL, "%, ", ci_instance_UL, "%]")`).

The single linked critique found appeared under the 'Recommended articles' sidebar of an article published in the Journal of Environmental Psychology webpage (for a depiction of how the article was linked, see Appendix F). This example was published as a letter[@Berto2020] and addressed concerns over incorrect citation of the critiquing author’s previous work and methodology used in the target article [@Hicks2020].

**2b) Articles Published in Any Psychology Journal**

No events of either linked or instances of post-publication critiques were identified in our random sample of 101 articles, resulting in a prevalence estimate of 0% (`r paste0(round(confidence_level*100), "% CI [", ci_instanceR_LL, "%, ", ci_instanceR_UL, "%]")`) for both classifications.

# Discussion
*Add 'Discussion' section*

# Methods
*Add 'Methods' section*

# Data Availability
The data used for this paper are available from the project’s OSF repository (link: [https://osf.io/8k7m4/](https://osf.io/8k7m4/))

# Code Availability
The analysis code for all results reported in this paper are available from the project’s OSF repository (link: [https://osf.io/8k7m4/](https://osf.io/8k7m4/))

# References