### 2024-04-11 analysis/resultsAnalysis.Rmd
Create resultsAnalysis.Rmd

---
title: "Results Analysis"
author: "Annie Whamond"
date: "2024-04-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r}
# Load libraries
library(here)
library(lsr)
library(tidyverse)
library(tidylog)
library(binom)
library(ggplot2)
library(RColorBrewer)
```

# Load data

```{r}
# Load preprocessed data
d_articles <- read_csv(here('data', 'processed', 'd_articles.csv'), show_col_types = F) # article data 
d_journals <- read_csv(here('data', 'processed', 'd_journals.csv'), show_col_types = F) # journal data
```

# STUDY ONE: JOURNAL POLICY

Remove excluded journals and create separate tibbles for prominent and randomly sampled journals
```{r}
dj_prominent <- d_journals %>%
  filter(sample_id != "random") %>%
  filter(is_empirical == TRUE)

dj_random <- d_journals %>%
  filter(sample_id != "prominent") %>%
  filter(is_empirical == TRUE)
```
## Journal characteristics
Descriptive statistics

Calculate JIF median and IQR for prominent journals
```{r}
# Prominent journal sample
prominentJIF <- dj_prominent %>% 
  filter(ppc_type == "A") %>%      # Keep only single entry for each journal
  summarise(promMedian_jif = round(median(jif),2), 
            promMin_jif = round(min(jif),2),
            promMax_jif = round(max(jif),2),
            promIQR_jif = round(IQR(jif),2)
  )

prominentJIF
```

Calculate JIF median and IQR for random journals
```{r}
randomJIF <- dj_random %>% 
  filter(ppc_type == "A") %>%
  filter(!is.na(jif)) %>%
  summarise(randMedian_jif = round(median(jif),2),
            randMin_jif = round(min(jif),2),
            randMax_jif = round(max(jif),2),
            randIQR_jif = round(IQR(jif),2)
  )

randomJIF
```

Count prominent journals listed as COPE members 
```{r}
prominentCOPE <- dj_prominent %>%
  filter(ppc_type == 'A') %>%
  filter(cope == TRUE) %>%
  summarise(count = n())

prominentCOPE
```

Count randomly sampled journals listed as COPE members 
```{r}
randomCOPE <- dj_random %>%
  filter(ppc_type == 'A') %>%
  filter(cope == TRUE) %>%
  summarise(count = n())

randomCOPE
```

Prominent journals Web Of Science categories
```{r}
prominentCategory <- dj_prominent %>%
  filter(ppc_type == 'A') %>%
  group_by(field) %>%
  summarise(count = n()) %>%
  ungroup()

prominentCategory
```
Random journals Web Of Science categories
```{r}
randomCategory <- dj_random %>%
  filter(ppc_type == 'A') %>%
  group_by(field) %>%
  summarise(count = n()) %>%
  ungroup()

randomCategory
```
## How many journals offer post-publication critique?

Inspect how frequently post-publication critique policies were stated for prominent journals 
```{r}
prominentPPC <- dj_prominent %>%
  filter(ppc_type == 'A') %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  ungroup()

prominentPPC
```
*Note: This may change once ARCHIVE ONLY coding is removed*

Inspect post-publication critique types at prominent journals
```{r}
prominentTypes <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  group_by(ppc_name) %>%
  summarise(count = n()) %>%
  ungroup()

prominentTypes
```

Inspect how frequently post-publication critique policies were stated for randomly sampled journals 
```{r}
randomPPC <- dj_random %>%
  filter(ppc_type == 'A') %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  ungroup()

randomPPC
```

Calculate confidence interval for explicitly stated post-publication critique policies among randomly sampled journals
```{r}

randomPPCyes <- dj_random %>%
  filter(ppc_type == 'A', !is.na(ppc_name)) %>%
  summarise(count = n())

# Calculate Wilson confidence intervals
sample_size <- 100
confidence_level <- 0.95

#CIs for 'Yes Explicit'
randomPPC_CI <- binom.confint(randomPPCyes, sample_size, method = "wilson", conf.level = confidence_level)

randomPPC_CI
```

Inspect post-publication critique types at randomly sampled journals
```{r}
randomTypes <- dj_random %>%
  filter(!is.na(ppc_name)) %>%
  group_by(ppc_name) %>%
  summarise(count = n()) %>%
  ungroup()

randomTypes
```